An efficient feedback calibration algorithm for direct imaging radio telescopes

### 2016-03-04 cal_demo/ms.tex
touching up cal paper

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/schematic.pdf}
\caption{The general data flow of the MOFF correlator, with a feedback calibration loop. A pixel from the (unsquared) image is tapped out and correlated against the input antenna electric field signals to form $\Cna$ coefficients (equation~\ref{eq:Cna}). These coefficients are then shipped off-chip to apply equations~\ref{eq:cal_solution}~and~\ref{eq:damping}, and any other processing to arrive at an updated estimate of the gains, $g^{(n)}_a$. The gray box shows operations which must be done at high speed (before averaging in time), which white boxes show operations which can be performed ``off-chip". ({\bf I have some suggestions to tweak this figure.})}
\label{fig:schematic}
\end{center}
\end{figure}

@@

Figure~\ref{fig:data_images} shows the improvement in the images due to our calibration. The left panel shows the uncalibrated image integrated over 51.2 ms, 29 kHz. Cyg~A is prominent near the center of the image, and Cas~A is also clearly visible in the upper right. The middle panel shows the image produced after calibration with identical integration time and bandwidth. The sidelobes throughout the image are significantly suppressed, and the galactic plane is much more evident, despite only modeling Cyg~A and Cas~A. We also note that the feature just to the right of Cyg~A is dimmer in the calibrated image, better matching the expected flux from the global sky model (GSM, \citealt{deo08}). We show the GSM\footnote{using the PyGSM module by D.~C.~Price to import the GSM, \textcolor{red}{insert reference}} modulated by two factors of the LWA power pattern in the right panel of figure~\ref{fig:data_images} for reference.

@@

The results of our simulations are shown in figure~\ref{fig:errors}. The gain errors from our EPICal method is shown with solid lines, while the dashed lines represent the visibility-based calibration. The thick lines represent the results of using a full sky model, and the thin lines are from the single point source model.  In the case of the full sky model, we see the errors trend downward with longer integration times, as expected. The EPICal errors are slightly higher than the visibility-based errors. This is not surprising as EPICal only uses the information in a single pixel, while the visibilities use all information from the full sky model. However, we see that with this perfect model the errors do behave like noise, and the same level of noise can be achieved with a longer integration time (or larger damping factor).

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/cal_paper_errors_complex_sky.pdf}
\caption{Gain estimate errors as a function of integration time, receiver noise (line color), and calibration method. The solid lines represent the error on the EPICal derived gain estimates, while the dashed lines represent the error on the visibility based estimates. The thicker lines were derived using a full sky model in the calibration loops, and the thinner lines used only a single point source model. In the full sky model case we see all methods trend down with longer integration time as expected. In the case of a single source model, the errors bottom out due to the flux in the measurements that is not modeled. In this regime the EPICal method performs slightly better than the visibility-based method due to effectively beamforming to the sky pixel where the incomplete model is most accurate.  
}
\label{fig:errors}
\end{center}
\end{figure}

When using the single point source model, all calibration methods trend downward until they reach an error floor due to the confusion sources that were not modeled. The EPICal and visibility-based solutions reach a similar floor, but EPICal achieves a marginally lower level. This can be attributed to the same reason EPICal underperformed with the full sky model: EPICal only used a single pixel in the sky to form its solutions. When using the full sky model, it was down weighting the information in the rest of the sky. But when the sky model only contains the bright point source, the single pixel used in EPICal is the location where this model is most accurate, effectively down weighting the pixels with incomplete sky model.

@@

There remain many avenues of future work before EPICal can be deployed on an instrument. Here we name a few potential considerations for further study. These topics can be thought of as extensions to the white ``estimate gains" box in figure~\ref{fig:schematic}, which can be performed at a much lower cadence than the actual correlations.

>>>>

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/schematic.pdf}
\caption{The general data flow of the MOFF correlator, with a feedback calibration loop. A pixel from the (unsquared) image is tapped out and correlated against the input antenna electric field signals to form $\Cna$ coefficients (equation~\ref{eq:Cna}). These coefficients are then shipped off-chip to apply equations~\ref{eq:cal_solution}~and~\ref{eq:damping}, and any other processing to arrive at an updated estimate of the gains, $g^{(n)}_a$. The gray box shows operations which must be done at high speed (before averaging in time), which white boxes show operations which can be performed ``off-chip".}
\label{fig:schematic}
\end{center}
\end{figure}

@@

Figure~\ref{fig:data_images} shows the improvement in the images due to our calibration. The left panel shows the uncalibrated image integrated over 51.2 ms, 29 kHz. Cyg~A is prominent near the center of the image, and Cas~A is also clearly visible in the upper right. The middle panel shows the image produced after calibration with identical integration time and bandwidth. The sidelobes throughout the image are significantly suppressed, and the galactic plane is much more evident, despite only modeling Cyg~A and Cas~A. We also note that the feature just to the right of Cyg~A is dimmer in the calibrated image, better matching the expected flux from the global sky model (GSM, \citealt{deo08}). We show the GSM\footnote{We use the PyGSM module by D.~C.~Price to import the GSM, https://github.com/telegraphic/PyGSM} modulated by two factors of the LWA power pattern in the right panel of figure~\ref{fig:data_images} for reference.

@@

The results of our simulations are shown in figure~\ref{fig:errors}. The gain errors from our EPICal method is shown with solid lines, while the dashed lines represent the visibility-based calibration. The thick lines represent the results of using a full sky model, and the thin lines are from the single point source model.  In the case of the full sky model, we see the errors trend downward with longer integration times, as expected. The EPICal errors are slightly higher than the visibility-based errors, on average about 23\% difference. This is not surprising as EPICal only uses the information in a single pixel, while the visibilities use all information from the full sky model. However, we see that with this perfect model the errors do behave like noise, and the same level of noise can be achieved with a longer integration time (or larger damping factor).

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figures/cal_paper_errors_complex_sky.pdf}
\caption{Gain estimate errors as a function of integration time, receiver noise (line color), and calibration method. The solid lines represent the error on the EPICal derived gain estimates, while the dashed lines represent the error on the visibility based estimates. The thicker lines were derived using a full sky model in the calibration loops, and the thinner lines used only a single point source model. In the full sky model case we see all methods trend down with longer integration time as expected. For a given integration time, EPICal derived errors are on average 23\% higher than those of visibility based calibration. In the case of a single source model, the errors bottom out due to the flux in the measurements that is not modeled. In this regime the EPICal method performs slightly better than the visibility-based method due to effectively beamforming to the sky pixel where the incomplete model is most accurate. On average EPICal errors are 5\% lower in this regime.
}
\label{fig:errors}
\end{center}
\end{figure}

When using the single point source model, all calibration methods trend downward until they reach an error floor due to the confusion sources that were not modeled. The EPICal and visibility-based solutions reach a similar floor, but EPICal achieves a marginally lower level (on average 5\% lower). This can be attributed to the same reason EPICal underperformed with the full sky model: EPICal only used a single pixel in the sky to form its solutions. When using the full sky model, it was down weighting the information in the rest of the sky. But when the sky model only contains the bright point source, the single pixel used in EPICal is the location where this model is most accurate, effectively down weighting the pixels with incomplete sky model.

@@

EPICal can be further improved through several extensions. Here we name a few potential considerations for further study. These topics can be thought of as extensions to the white ``estimate gains" box in figure~\ref{fig:schematic}, which can be performed at a much lower cadence than the actual correlations.
